### 2019-10-13 Moral_Communities_and_Crime_v1.Rmd
Completed revisions to Background section and abstract.

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Aye B. Cee
  address: A Department, University of XXX
  email: abc@example.edu

- name: Dee E. Eff
  address: A Department, University of XXX
  email: def@example.edu

- name: Gee H. Eye
  address: A Department, University of XXX
  email: ghi@example.edu

- name: Jay K. Elle
  address: Some School, University of WWW
  email: jkl@school.edu

keywords:
- Moral communities
- Crime
- Places of worship
- Point pattern analysis
- Intensity

abstract: |
  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries. With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities. Moral communities, a concept closely related to informal social controls, are of interest in countries and regions where formal controls are weak and ineffective. The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship. Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns. Analysis is condcuted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets). The results suggest that Catholic and Spiritist places of worship do not project moral communities geographically more than, say, ice cream shops. The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher than for any of our other referential events.

bibliography: bibliography.bib
output: rticles::springer_article
---

<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following file is needed: `Moral_Communities_and_Crime.RData`. For reference: this data file was created using R notebook `00. Data Preprocessing.Rmd`. The initial analysis was done in R notebook `01. Analysis.Rmd`. Neither of these two files are needed for this document. -->

```{r load-packages, include=FALSE}

# Load packages used in this document

library(tidyverse)
library(kableExtra)
library(sf)
library(spatstat)
library(maptools)
library(ggthemes)
library(raster)
library(fasterize)
```

```{r load-data, include=FALSE}

# To run this notebook, a file `Moral_Communities_and_Crime.RData` is needed. This file contains the following objects:
# 
# - events_recife.sf: a simple features object with points for violent crimes, churches (Catholic, Evangelical, Spiritist) and commercial establishments (ice cream shops, pharmacies, supermarkets) in the city of Recife.
# 
# - recife.sf: a simple features object with the boundaries of the city of Recife
# 
# - setores.sf: a simple features object with setores (census tracts) and selected socio-economic and demographic information for the city of Recife, including:
#   - Population density in persons per sq.km
#   - Median income in setor
#   - Proportion of population in setor who are unemployed
#   - Proportion of households with a single mother
#   - Proportion of households with a young single mother, a woman between the ages of 15 and 25
#   - Proportion of population in poverty
# 
# Read the RData file:

load("Moral_Communities_and_Crime.RData")
```

```{r data-preparation, include=FALSE, cache=TRUE}

# In preparation for doing point pattern analysis, convert the `recife.sf` object to an `owin` object for use in `spatstat`:

recife.owin <- recife.sf %>% as("Spatial") %>% as("owin")
plot(recife.owin)

# Convert the events to a `ppp` object:

events_recife.ppp <- events_recife.sf %>% 
  st_coordinates() %>% 
  data.frame() %>%
  mutate(Type = events_recife.sf$Type) %>%
  as.ppp(W = recife.owin)

#Split the point pattern into each type of event: crime, Catholic place of worship, Evangelical place of worship, Spiritist place of worship, ice cream shop, pharmacy, and supermarket

events_recife_list <- split(events_recife.ppp)
```

```{r socio-economic-deprivation, include=FALSE, cache=TRUE}

# Here an indicator of socio-economic deprivation is obtained using principal component analysis

# PCA is implemented in R as `prcomp`:
setores.pca <- setores.sf %>% 
  st_drop_geometry() %>%
  dplyr::select(median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
  prcomp(center = TRUE, scale = TRUE)

# The loadings of the pca are as follows:
setores.pca$rotation

# Socio-economic deprivation is a combination of high poverty, high proportion of single mothers, especially young women between the ages of 15-25, high proportion of unemployed people, and low median income.

# Various diagnostics of the results are possible, such as the variance:

std_dev <- setores.pca$sdev

# Compute variance

pr_var <- std_dev^2

# Check variance of the components

pr_var[1:5]

# And the proportion of the variance:

pr_var/sum(pr_var)

# Once that there is principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:

setores.sf$SED <- setores.pca$x[,1]

# Normalize the SDE factor (with the addition of a small correction; the `rhohat` function does not like values <= 0):

setores.sf <- setores.sf %>% mutate(SED = (SED - min(SED) + 0.1)/(max(SED) + 0.1 - min(SED)))
```

# Introduction {#intro}

Violent crime is a widespread phenomenon with negative impacts on many spheres of life and society. Although the rate of homicides worldwide has grown at a slower rate than the population in the past few decades, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive]. Furthermore, variations in the prevalence of violent crime tend to be extremely uneven. The Americas, with a population of approximately 793.8 million people in 2019 (or about 10.3% of the world population), accounted for approximately 173,000 homicides in 2017, or approximately 37.3% of all homicides in the world. Within the Americas, Brazil (along with Venezuela, Colombia, and Mexico) is one of the largest countries in the region with high homicide rates. And there are notable variations at the sub-national and intra-metropolitan levels as well. In fact, understanding the variations in the prevalence of violent crime is recognized as key to achieving policy goals:

>> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so
achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p. 35]

This concern is not iddle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals. Violent crime, furthermore, tends to be more accute in those regions most in need of development. On the one hand, violent crime represents an economic and social drag [@Becker2017analise]. The impacts are multifaceted, since they affect people's well-being, either through the loss of human life, mental health, and limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development]. In turn, these negative effects combine to create the unfortunate conditions that tend to breed criminal behavior, thus creating a vicious cycle of economic disadvantage and violence [@Unodc2019development]. It is not surprising, given these pecularities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].

Added to the scenario, formal social control institutions in countries like Brazil leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up reducing the deterrent power of the justice system. Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial]. This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborbood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].

In order to understand the factors that can potentially help to deter crime, beyond formal control, it is important to identify empirical regularities. Criminological factors include concrete elements (such as te presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the soci costs of deviant behavior as well as family supervision. Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets [e.g., @Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.

With the above considerations in mind, the objective of this paper is to investigate whether and how places of workship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension. The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast. Recife is a large and important metropolitan area in a historically poor region, and afflicted by high levels of violent crime, having the dubious honor of being one of five state capitals with the highest rates of homicide in the period under study [@Menezes2013spatial]. The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments that serve as controls, on the other. Dissagregated data allows us to analyze the relevant point patterns at a very high level of resolution.

After this introductory section, the rest of this paper is structured as follows: Section \ref{background} reviews some relevant theoretical perspectives on religion and crime; then Section \ref{methods} presents the empirical strategy used in the work. Section \ref{case} presents the context as well as the data used in the study; Section \ref{results} presents the analysis, whereas the results are discussed in Section \ref{discussion}. Finally, some concluding remarks are offered in Section \ref{conclusions}.

# Background: Theoretical Perspectives on Religion and Crime {#background}

Why do humans behave morally? For millennia, it has been the role of religion to provide the basic tenets of morally acceptable behavior: thou shalt not kill et al., and their equivalent in many world religions [e.g., @Donovan1986different]. Enforcement of such tenets implies different mechanisms, including _sin_, _haram_, _karma_, and _tapu_, with punishment delivered by hellfire and exile to _Jahannam_ or _Gehenna_, to mention just some choice places of torment. In addition to acting as _de facto_ social policy for much of history, many of these religious tenets still have the force of law in many cultures and regions - and even where they do not, they are held by some researchers and policy experts to be helpful complements to reduce crime in any case [e.g., @Durrant2017religion; @Johnson2011more]. The hypothesis that religion can act as a factor that deters and reduces criminal behavior, therefore, has prompted the scientific study of the effectiveness of religion on moral behavior [@Hoffmann2015religion].

An early effort to theorize the effect of religion on criminal behavior was the _hellfire hypothesis_ of Hirschi and Stark [-@Hirschi1969hellfire]. The focus of this hypothesis is the threat of extratemporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager). Accordingly, Hirsch and Stark [-@Hirschi1969hellfire] posit that negative correlations between crime and religion are a consequence of a sense of commitment with normative values, a commitment that is ritually reinforced in a regular fashion (e.g., the liturgical rite of peace in Catholic Mass). Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity], for instance, note that by attributing to divinity (or some other supra-terrenal entity) the supreme force of punishment, religion helps to build a sense of obduracy, or "hardening" against temptation.

Despite its intuitive appeal, research has not provided much direct evidence for the hypothesis of hellfire [@Hoffmann2015religion, p. 1]. In a bizarre twist, even mainstream doctrines (such as absolution) may in fact encourage beliefs that neutralize the fear of _terrenal_ punishment, and thus turn out to be criminogenic [e.g., @Topalli2013god]. More deviant cases can even recast criminal behavior as a form of spiritual insurgency [@Chesnut2017devoted]. Counterexamples like this notwithstanding, researchers remain open to the possibility that "religion may still serve as a social control mechanism by encouraging conventional beliefs, monitoring behaviors, enhancing family attachments, or providing conventional activities" [@Hoffmann2015religion, p. 1]. In this way, the hypothesis of _moral communities_ [@Stark1996religion] recognizes that social integration is essential for increasing social control, thus reducing the practice of behavior that is not in accordance with current norms. Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective” [@Reiss1951delinquency, p. 196; cited in @Groff2015informal, p. 91].

The hypothesis of moral communities has over the years been used to examine a variety of outcomes of interest. Recent examples include Stroope and Baker's [-@Stroope2018moral] exploration of religiosity and self-rated health; Davignon and Thomson [-@Davignon2015christian] with their research on institutional context and the religiosity of students; and the study of religion as a source of trust of Traunmuller [-@Traunmuller2011moral]. This is in addition, of course, to numerous studies on criminal behavior such as Eitle [-@Eitle2011religion], an author who explored the deterrence power of religion on gambling; Lee and Bartkowski's [-@Lee2004love] investigation of juvenile homicide in rural areas; and the research of Regnerus [-@Regnerus2003moral] on adolescent delinquency. The hypothesis of moral communities as a form of informal social control, on the other hand, has been less studied from a geographical perspective, and it is only recently that has attracted the attention of researchers. Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plaussibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county. In this way, Nie and Yang [-@Nie2019smoking] remark on a recent paper on the lack of research conducted to study how the religious context of a geographical area (e.g., a county) may influence (youth smoking) behavior (p. 2). This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory are thought to occur at relatively small geographical scales.

The preceding review makes clear the need for research at various geographical scales. From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood. In what follows, we propose to adopt a much more disaggregated approach by considering the intensity of criminal events with respect to places of worship. This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.

# Empirical Strategy {#methods}

Two different techniques are deployed in this paper to study the spatial distribution of criminal events, and their possible relationship to the location of places of worship. Firstly, the F-function is a nearest neighbor technique that is used to summarize the distance between a set of points and their nearest events. This technique has found extensive use in spatial epidemiology (see Gatrell et al., 1996) and has also been used in spatial criminology (see, inter alia, Craglia et al., 2000; Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001). The second technique is the scan statistic. This method has also been extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996) and has found use in spatial criminology as well (see He et al., 2017a; Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011). Use of these two techniques is meant to provide insights with respect to 1) the location of violent criminal events with respect to various types of facilities; and 2) geographically identify places of worship associated with clusters of criminal events.
These two techniques are briefly reviewed next.

## Quadrat Analysis

Words go here.

## Relative Intensity

Words go here.

# Case Study {#case}

## Context {#context}

The study is of the city of Recife, the capital of the state of Pernambuco in the Northeast region of Brazil. With a population of 1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions. However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].

In the current context, the term "violent crime" is an umbrella for several forms of infractions to the penal code. Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (VLIC; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death. The data on CVLI were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.

```{r descriptive-statistics, echo=FALSE}
descriptive_stats <- data.frame(Attribute = c("Man",
                                              "Woman",
                                              "Murder",
                                              "Robbery",
                                              "Body injury followed by death",
                                              "Black and White",
                                              "Yellow and White",
                                              "Not Reported",
                                              "1 to 12 years",
                                              "13 to 17 years",
                                              "18 to 30 years",
                                              "31 to 65 years",
                                              "65 years and older",
                                              "Not Reported",
                                              "Firearm",
                                              "Other"),
                                Percentage = c(91.35,
                                               8.65,
                                               97.82,
                                               2.00,
                                               0.18,
                                               95.11,
                                               1.27,
                                               3.62,
                                               0.18,
                                               11.31,
                                               61.70,
                                               25.84,
                                               0.73,
                                               0.24,
                                               87.51,
                                               12.49))

kable(descriptive_stats, 
      "latex", 
      booktabs = TRUE,
      digits = 3,
      caption = "\\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009",
      col.names = c("", "Percentage")) %>%
  kable_styling() %>%
  pack_rows("Gender of Victim", 1, 2) %>%
  pack_rows("Type of Crime", 3, 5) %>%
  pack_rows("Ethnicity of Victim", 6, 8) %>%
  pack_rows("Age of Victim", 9, 14) %>%
  pack_rows("Weapon Used", 15, 16)
```

## Data {#data}

The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim. The database used in this study comprises the period from July 1, 2008 to June 30, 2010. Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of CVLI in the period analyzed were men, while approximately 98% of violent crimes were homicides. In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime. Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms.
Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts. This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census. The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination. Georeferencing was used to geolocate each place of worship. In this way, a total of 1,719 places of worship were geolocated in the city of Recife.
Figure 1 (Panel a) shows the spatial distribution of the 1,657 CVLI crimes that occurred in the city of Recife between 2008.2 and 2010.1. Panel b) of the figure shows the spatial distribution of places of worship in the region.

```{r plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife"}

#We can experiment with plotting the events:

ggplot(data = events_recife.sf %>% filter(Type == "Violent Crime")) + 
  geom_sf(color = "red") +
  geom_sf(data = recife.sf, fill = NA) + 
  theme_tufte()
```

In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship. As discussed above, the idea is to identify points of reference that can be used as controls, having a neutral morality profile. For the sake of the present study, we selected pharmacies, ice cream shops, and bakeries to construct our control group. These three types of establishments comply with the criteria of being morally neutral and having a spatial distribution commensurate with places of worship.
Figure 2 shows the spatial distribution of the control establishments, namely pharmacies, bakeries, and ice cream shops in the city of Recife. Note the similarity between the maps. As expected, there are differences in the number of points, but the locations of cases and controls are quite similar.

```{r plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife"}

#We can experiment with plotting the events:

ggplot(data = events_recife.sf %>% filter(Type != "Violent Crime")) + 
  geom_sf(aes(color = Type, fill = Type)) +
  geom_sf(data = recife.sf, fill = NA) + 
  facet_wrap(~ Type, nrow = 2) +
  scale_fill_viridis_d(option = "viridis", direction = 1) +
  scale_color_viridis_d(option = "viridis", direction = 1) +
  theme_tufte()  +
  theme(axis.text.x = element_text(angle = 90))

```

```{r summary-point-patterns, include=FALSE}

# Extract the intensity of each point pattern for the study area

summary_crime <- summary(events_recife_list$`Violent Crime`)
summary_catholic <- summary(events_recife_list$Catholic)
summary_evangelic <- summary(events_recife_list$Evangelical)
summary_spiritist <- summary(events_recife_list$Spiritist)
summary_ice_cream <- summary(events_recife_list$`Ice Cream Shop`)
summary_pharmacy <- summary(events_recife_list$Pharmacy)
summary_supermarket <- summary(events_recife_list$Supermarket)
```

```{r plot-sed, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
ggplot(data = setores.sf) + 
  geom_sf(aes(fill = SED),
          color = NA) +
  scale_fill_viridis_c(option = "plasma", direction = 1, values = c(0, 0.05, 0.2, 0.3, 0.4, 1)) +
  theme_tufte()
```

# Analysis and results {#results}

## Quadrat Analysis

```{r rasterize-sed, include=FALSE}

# Rasterize and convert to image:

# First, we need a base raster; this will be the boundaries of the study area. After some trial and error with different values of number of rows and columns, these values create pixels that are as close to squares as possible, while also being relatively small

recife.r <- raster(recife.sf, nrows = 400, ncols = 276)

# The raster `recife.r` is the template for rasterizing other items, in this case socio-economic deprivation. Use package `fasterize` to rasterize an `sf` object.

sed.r <- fasterize(setores.sf, recife.r, field = "SED")

# Once that the object is a raster, it can be converted to a `spatstat` `img` class object:
sed.img <- as.im(sed.r)
```

```{r quadrat-analysis-preliminaries, include=FALSE}

# To do quadrat analysis by different levels of SED we need the quintiles of this variable (other criteria could be used to slice the distribution, e.g., standard deviations, but this one is intuitive and simple)

sed5 <- quantile(sed.img, c(0, 0.2, 0.4, 0.6, 0.8, 1))

# Next we can cut the `sed.im` using the quintiles:

sed_cut5.img <- cut(sed.img, breaks = sed5, include.lowest = TRUE, labels = c(1:5))
```

```{r tesselate-sed5, include=FALSE}

# To do quadrat analysis we need to tesellate the cut image:

V <- tess(image = sed_cut5.img)

# This object is used for the analysis, but it is not very neat for plotting purposes. Lets convert instead to an `sf` object. First, extract and expand the coordinates:

coords_V <- expand.grid(X = V$image$xcol, Y = V$image$yrow)

# Next, extract the values of the cells and convert to a vector:

sed5_V <- as.vector(t(V$image$v))

# Join the coordinates with the values in a dataframe and drop all NAs, before converting to `sf` and adding coordinate referencing system information (crs) to ensure that its projection information is there and compatible with the rest of `sf` objects (i.e., recife.sf, setores.sf, events_recife.sf):

sed5.sf <- data.frame(coords_V, sed5_V) %>% 
  drop_na() %>% transmute(X, Y, SED = factor(sed5_V, 
                                              levels = c(1:5),
                                              labels = c("Low", 
                                                         "Medium Low", 
                                                         "Medium", 
                                                         "Medium High", 
                                                         "High"))) %>%
  st_as_sf(coords = c("X", "Y")) %>%
  st_set_crs(value = st_crs(recife.sf))
```

```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as Quintiles" }
ggplot(data = sed5.sf) + 
  geom_sf(aes(fill = SED, color = SED)) +
  geom_sf(data = recife.sf, fill = NA) +
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_color_viridis_d(option = "plasma", direction = 1) +
  guides(fill = guide_legend(reverse = TRUE),
         color = guide_legend(reverse = TRUE)) +
  theme_tufte()
```

Calculate the intensity of other ppp by quadrat:
```{r quadrat-analysis, include=FALSE}

# Calculate the intensity of the various point patterns by quadrat:

# Crime

crime5 <- quadratcount(events_recife_list$`Violent Crime`, tess = V)
crime5_intensity <- intensity(crime5)

# Catholic

catholic5 <- quadratcount(events_recife_list$Catholic, tess = V)
catholic5_intensity <- intensity(catholic5)

# Evangelical

evangelic5 <- quadratcount(events_recife_list$Evangelical, tess = V)
evangelic5_intensity <- intensity(evangelic5)

# Spiritist

spiritist5 <- quadratcount(events_recife_list$Spiritist, tess = V)
spiritist5_intensity <- intensity(spiritist5)

# Ice cream

ice_cream5 <- quadratcount(events_recife_list$`Ice Cream Shop`, tess = V)
ice_cream5_intensity <- intensity(ice_cream5)

# Pharmacies

pharmacies5 <- quadratcount(events_recife_list$Pharmacy, tess = V)
pharmacies5_intensity <- intensity(pharmacies5)

# Supermarkets

super_markets5 <- quadratcount(events_recife_list$Supermarket, tess = V)
supermarkets5_intensity <- intensity(super_markets5)
```

```{r instensity-dataframe, include=FALSE}

# Collect in a data.frame for convenience of plotting:

intensity_all <- data.frame(SED = rep(c("Low", 
                                        "Middle Low", 
                                        "Middle", 
                                        "Middle High", 
                                        "High"), 
                                      times = 7),
                            Type = rep(c("Violent Crime", 
                                         "Catholic",
                                         "Evangelical",
                                         "Spiritist",
                                         "Ice Cream",
                                         "Pharmacies",
                                         "Super Markets"), 
                                       each = 5),
                            Intensity = c(crime5_intensity,
                                          catholic5_intensity,
                                          evangelic5_intensity,
                                          spiritist5_intensity,
                                          ice_cream5_intensity,
                                          pharmacies5_intensity,
                                          supermarkets5_intensity),
                            Global_Intensity = rep(c(summary_crime$intensity,
                                                     summary_catholic$intensity,
                                                     summary_evangelic$intensity,
                                                     summary_spiritist$intensity,
                                                     summary_ice_cream$intensity,
                                                     summary_pharmacy$intensity,
                                                     summary_supermarket$intensity),
                                                   each = 5)) %>%
  mutate(SED = factor(SED, levels = c("Low", 
                                        "Middle Low", 
                                        "Middle", 
                                        "Middle High", 
                                        "High"),
                      ordered = TRUE),
         Type = factor(Type, levels = c("Violent Crime", 
                                        "Catholic",
                                        "Evangelical",
                                        "Spiritist",
                                        "Ice Cream",
                                        "Pharmacies",
                                        "Super Markets")))
```

```{r plot-crime-quadrat, echo=FALSE, fig.width=3, fig.asp=1 , cache = TRUE, fig.cap="\\label{fig:plot-crime-quadrat}Intensity of crime by level of Socio-Economic Deprivation; the dotted line indicates the global intensity of crime"}

#Plot the intensity of crime:

ggplot(data = intensity_all %>% filter(Type == "Violent Crime"), 
       aes(x = SED, y = Intensity, fill = Type)) + 
  geom_col(fill = "red") +
  xlab("Socio-Economic Deprivation") +
  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
  #facet_wrap(Type ~.) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r plot-events-quadrat, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-events-quadrat}Intensity of places of worship and commercial establishments by level of Socio-Economic Deprivation; the dotted lines indicates the respective global intensities"}

#Plot the intensity of point patterns:

ggplot(data = intensity_all %>% filter(Type != "Violent Crime"), 
       aes(x = SED, y = Intensity, fill = Type)) + 
  geom_col() +
  xlab("Socio-Economic Deprivation") +
  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
  scale_fill_viridis_d(option = "viridis", direction = 1) +
  scale_color_viridis_d(option = "viridis", direction = 1) +
  facet_wrap(Type ~.) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

## Relative Intensity

```{r distance-functions, include=FALSE, cache = TRUE}

# Estimation of the relative intensity (see Baddeley et al. (2015) Spatial Point Pattern, pp. 179-183).

# This needs that we convert the point patterns for places of worship to distance functions. This way, the covariates for the relative distribution estimate are distance estimates to places of worship. The function `distfun` is used for this:

z_catholic <- distfun(events_recife_list$Catholic)
z_evangelical <- distfun(events_recife_list$Evangelical)
z_spiritist <- distfun(events_recife_list$Spiritist)
z_ice_cream <- distfun(events_recife_list$`Ice Cream Shop`)
z_pharmacies <- distfun(events_recife_list$Pharmacy)
z_super_markets <- distfun(events_recife_list$Supermarket)
```

# Discussion {#discussion}

An opposing viewpoint is offered by the theory of conservative Protestants [@Ellison2003enduring]. This theory argues that religion is not always beneficial in order to guarantee social order, and consequently to result in lower crime rates. This theory suggested that, conservative Protestants are more tolerate with violent behavior when associated with defense of honor, family, property, or woman. However, high rates are not expected for all types of crimes in densely populated areas of conservative Protestants. But what justifies this behavior? Ellison et al. [-@Ellison2003enduring] argue that this tolerance stems from a religious view of the world. More clearly, punishment is seen as a legitimate form of retribution whose explanation can be attributed to the literal interpretation of the Bible. In the specific context of Brazil, it is important to consider that the interpretations and meanings attributed to urban violence during the last two decades in Brazil often had religious figures as protagonists [@Birman2012violencia].

# Conclusions {#conclusions}

The theory of conservative Protestants, argues that conservative Protestants are more likely to tolerate violent behavior of a defensive or punitive character, especially when associated to defense of honor, family, property, or woman. Our objective in the study is to explore the question whether moral communities have a spatial dimension. To this end we deploy an array of spatial analytical tools to investigate potential patterns of clustering of violent crime and places of worship, using as well other kinds of urban facilities to act as controls. The case study is the city of Recife, in the northeast of Brazil, and analysis is based on geographically detailed datasets of criminal events as well as several types of urban facilities.

# References