### 2019-10-13 Moral_Communities_and_Crime_v1.tex

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\title{Do moral communities have a spatial dimension? A spatial exploratory
analysis of places of worship and violent crime in the city of Recife,
Brazil \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

    \titlerunning{Moral communities and crime}

\author{  Aye B. Cee \and  Dee E. Eff \and  Gee H. Eye \and  Jay K. Elle \and  }

\institute{
        Aye B. Cee \at
     A Department, University of XXX \\
     \email{\href{mailto:abc@example.edu}{\nolinkurl{abc@example.edu}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Dee E. Eff \at
     A Department, University of XXX \\
     \email{\href{mailto:def@example.edu}{\nolinkurl{def@example.edu}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Gee H. Eye \at
     A Department, University of XXX \\
     \email{\href{mailto:ghi@example.edu}{\nolinkurl{ghi@example.edu}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Jay K. Elle \at
     Some School, University of WWW \\
     \email{\href{mailto:jkl@school.edu}{\nolinkurl{jkl@school.edu}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Religious tenets of the type ``thou shalt not kill'' and their
equivalents in many world religions have functioned as \emph{de facto}
social policy for determining appropriate and acceptable behavior over
the centuries. With the advent of the scientific study of religion,
there has been a growth of interest in the role of religions to operate
as moral communities. Moral communities, a concept closely related to
informal social controls, are of interest in countries and regions where
formal controls are weak and ineffective. The objective of this paper is
to present a spatial analysis of Violent and Intentional Crime in the
city of Recife in Brazil, with a focus on the possible interactions
between criminal events and places of worship. Previous research into
moral communities has advocated the need for analysis at different
scales, and this analysis contributes to the literature by using
micro-level data and appropriate spatial analytical tools for spatial
point patterns. Analysis is condcuted using three different types of
places of worship (Catholic, Evangelical, and Spiritist) and three types
of business establishments as controls (ice cream shops, pharmacies, and
supermarkets). The results suggest that Catholic and Spiritist places of
worship do not project moral communities geographically more than, say,
ice cream shops. The intensity of criminal events in the proximity of
Evangelical places of worship, in contrast, is markedly higher than for
any of our other referential events.
\\
\keywords{
        Moral communities \and
        Crime \and
        Places of worship \and
        Point pattern analysis \and
        Intensity \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

Violent crime is a widespread phenomenon with negative impacts on many
spheres of life and society. Although the rate of homicides worldwide
has grown at a slower rate than the population in the past few decades,
the number of people killed in homicides still increased from 362,000 in
1990 to 464,000 in 2017 (United Nations Office on Drugs and Crime
2019a). Furthermore, variations in the prevalence of violent crime tend
to be extremely uneven. The Americas, with a population of approximately
793.8 million people in 2019 (or about 10.3\% of the world population),
accounted for approximately 173,000 homicides in 2017, or approximately
37.3\% of all homicides in the world. Within the Americas, Brazil (along
with Venezuela, Colombia, and Mexico) is one of the largest countries in
the region with high homicide rates. And there are notable variations at
the sub-national and intra-metropolitan levels as well. In fact,
understanding the variations in the prevalence of violent crime is
recognized as key to achieving policy goals:

\begin{quote}
\begin{quote}
High levels of homicidal violence are concentrated in geographic and
demographic ``pockets'', so achieving target 16.1 of the Sustainable
Development Goals requires interventions within the specific regions,
countries, communities and population groups that are most at risk
(United Nations Office on Drugs and Crime 2019a, 35)
\end{quote}
\end{quote}

This concern is not iddle: the endemic malady of violent crime
represents an important factor that threatens to derail sustainable
development goals. Violent crime, furthermore, tends to be more accute
in those regions most in need of development. On the one hand, violent
crime represents an economic and social drag (Becker and Kassouf 2017).
The impacts are multifaceted, since they affect people's well-being,
either through the loss of human life, mental health, and limitations in
the right to public spaces (Doran and Burgess 2011), or through
disturbances in schooling and academic achievement (United Nations
Office on Drugs and Crime 2019b). In turn, these negative effects
combine to create the unfortunate conditions that tend to breed criminal
behavior, thus creating a vicious cycle of economic disadvantage and
violence (United Nations Office on Drugs and Crime 2019b). It is not
surprising, given these pecularities, that researchers have joined calls
for increased attention to the study of the patterns of violent crime in
Low and Middle Income Countries (LMIC; Murray, Cerqueira, and Kahn
2013).

Added to the scenario, formal social control institutions in countries
like Brazil leave something (or much) to be desired: while in practice
they should inhibit criminal behavior, they suffer from deep
deficiencies that end up reducing the deterrent power of the justice
system. Serious institutional problems include the inefficiency of the
police, the lack of national legislation, the glacial pace of judicial
processes, and the weak situation of the prison system in the country
(Menezes et al. 2013). This makes it even more urgent to understand the
role of \emph{informal social controls}, i.e., the ability of community
organizations such as schools, clubs, and neighborbood associations to
suppress crime by strengthening the capabilities of neighbors to control
inappropriate behavior (Groff 2015).

In order to understand the factors that can potentially help to deter
crime, beyond formal control, it is important to identify empirical
regularities. Criminological factors include concrete elements (such as
te presence of arms or drugs, or elements of the built environment), and
also figurative factors, which include the soci costs of deviant
behavior as well as family supervision. Accordingly, a number of studies
have investigated various aspects of the environment and neighborhood
design (e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu
2017; Loukaitou-Sideris et al. 2001), whereas other studies have focused
on exposure to environmental attributes that signal weakened norms, such
as liquor and tobacco outlets (e.g., Brower and Carroll 2007; Deryol et
al. 2016; Lipton et al. 2008; Quick, Law, and Luan 2017).

Yet another fruitful avenue for research, and one that has only recently
begun to be explored, is the presence of environmental attributes that
can help to reinforce moral norms, such as schools and churches (e.g.,
Abdullah et al. 2018; Davignon and Thomson 2015; Furr-Holden et al.
2010; Traunmuller 2011). It is thus that in a recent paper, Warner and
Konkel (2019) note that the role of places of worship, as distinct
entities from the members of the congregations, have received less
attention in empirical and theoretical research for their deterrence
potential. The role of these institutions might be particularly
important in places where formal state institutions lack the means or
the will to enforce norms - as is the case in Brazil.

With the above considerations in mind, the objective of this paper is to
investigate whether and how places of workship correlate spatially with
criminal events, or in other words, to investigate whether their signals
as moral communities have a discernible spatial dimension. The case
study is the city of Recife in the state of Pernambuco, in Brazil's
Northeast. Recife is a large and important metropolitan area in a
historically poor region, and afflicted by high levels of violent crime,
having the dubious honor of being one of five state capitals with the
highest rates of homicide in the period under study (Menezes et al.
2013). The empirical strategy is to use spatial analysis to explore the
potential geographical relationships between violent criminal events, on
the one hand, and places of worship and a selection of commercial
establishments that serve as controls, on the other. Dissagregated data
allows us to analyze the relevant point patterns at a very high level of
resolution.

After this introductory section, the rest of this paper is structured as
follows: Section \ref{background} reviews some relevant theoretical
perspectives on religion and crime; then Section \ref{methods} presents
the empirical strategy used in the work. Section \ref{case} presents the
context as well as the data used in the study; Section \ref{results}
presents the analysis, whereas the results are discussed in Section
\ref{discussion}. Finally, some concluding remarks are offered in
Section \ref{conclusions}.

\hypertarget{background}{%
\section{Background: Theoretical Perspectives on Religion and
Crime}\label{background}}

Why do humans behave morally? For millennia, it has been the role of
religion to provide the basic tenets of morally acceptable behavior:
thou shalt not kill et al., and their equivalent in many world religions
(e.g., Donovan 1986). Enforcement of such tenets implies different
mechanisms, including \emph{sin}, \emph{haram}, \emph{karma}, and
\emph{tapu}, with punishment delivered by hellfire and exile to
\emph{Jahannam} or \emph{Gehenna}, to mention just some choice places of
torment. In addition to acting as \emph{de facto} social policy for much
of history, many of these religious tenets still have the force of law
in many cultures and regions - and even where they do not, they are held
by some researchers and policy experts to be helpful complements to
reduce crime in any case (e.g., Durrant and Poppelwell 2017; Johnson
2011). The hypothesis that religion can act as a factor that deters and
reduces criminal behavior, therefore, has prompted the scientific study
of the effectiveness of religion on moral behavior (Hoffmann 2015).

An early effort to theorize the effect of religion on criminal behavior
was the \emph{hellfire hypothesis} of Hirschi and Stark (1969). The
focus of this hypothesis is the threat of extratemporal (and possibly
eternal) punishment, and how this threat can deter believers from
commandment-breaking (also see Pascal's wager). Accordingly, Hirsch and
Stark (1969) posit that negative correlations between crime and religion
are a consequence of a sense of commitment with normative values, a
commitment that is ritually reinforced in a regular fashion (e.g., the
liturgical rite of peace in Catholic Mass). Rohrbaugh and Jessor (1975),
for instance, note that by attributing to divinity (or some other
supra-terrenal entity) the supreme force of punishment, religion helps
to build a sense of obduracy, or ``hardening'' against temptation.

Despite its intuitive appeal, research has not provided much direct
evidence for the hypothesis of hellfire (Hoffmann 2015, 1). In a bizarre
twist, even mainstream doctrines (such as absolution) may in fact
encourage beliefs that neutralize the fear of \emph{terrenal}
punishment, and thus turn out to be criminogenic (e.g., Topalli,
Brezina, and Bernhardt 2013). More deviant cases can even recast
criminal behavior as a form of spiritual insurgency (Chesnut 2017).
Counterexamples like this notwithstanding, researchers remain open to
the possibility that ``religion may still serve as a social control
mechanism by encouraging conventional beliefs, monitoring behaviors,
enhancing family attachments, or providing conventional activities''
(Hoffmann 2015, 1). In this way, the hypothesis of \emph{moral
communities} (Stark 1996) recognizes that social integration is
essential for increasing social control, thus reducing the practice of
behavior that is not in accordance with current norms. Rohrbaugh and
Jessor (1975) also emphasize that religion acts as social control since
it defines what is an appropriate attitude according to moral values,
thus making moral communities a close relative of the concept of
\emph{informal social control}, defined as ``the ability of social
groups or institutions to make norms or rules effective'' (Reiss 1951,
196; cited in Groff 2015, 91).

The hypothesis of moral communities has over the years been used to
examine a variety of outcomes of interest. Recent examples include
Stroope and Baker's (2018) exploration of religiosity and self-rated
health; Davignon and Thomson (2015) with their research on institutional
context and the religiosity of students; and the study of religion as a
source of trust of Traunmuller (2011). This is in addition, of course,
to numerous studies on criminal behavior such as Eitle (2011), an author
who explored the deterrence power of religion on gambling; Lee and
Bartkowski's (2004) investigation of juvenile homicide in rural areas;
and the research of Regnerus (2003) on adolescent delinquency. The
hypothesis of moral communities as a form of informal social control, on
the other hand, has been less studied from a geographical perspective,
and it is only recently that has attracted the attention of researchers.
Groff (2015), for instance, discusses informal social control as a
phenomenon that can plaussibly operate at different geographical scales,
from the level of the home and family, through the street block and
neighborhood, possibly to other scales such as the county. In this way,
Nie and Yang (2019) remark on a recent paper on the lack of research
conducted to study how the religious context of a geographical area
(e.g., a county) may influence (youth smoking) behavior (p.~2). This
point is echoed by Warner and Konkel (2019), who moreover recommend the
use of even smaller units of analysis (e.g., Census Block Groups), in an
effort to reduce aggregation bias (see Hipp 2007), but more importantly
because the processes defined by social disorganization theory are
thought to occur at relatively small geographical scales.

The preceding review makes clear the need for research at various
geographical scales. From a theoretical perspective, the social
mechanisms that underpin the hypothesis of moral communities and
informal social control can happen at various geographical scales, some
of which previous research has already addressed: the work of Nie and
Yang (2019), for one, highlights the role of processes detectable at a
relatively high level of aggregation, whereas Warner and Konkel (2019),
looking specifically at the effect of places of worship, bring the study
closer to the level of the neighborhood. In what follows, we propose to
adopt a much more disaggregated approach by considering the intensity of
criminal events with respect to places of worship. This requires
micro-level data both of crimes and places of worship, and the use of
appropriate spatial analytical tools for point patterns, as discussed
next.

\hypertarget{methods}{%
\section{Empirical Strategy}\label{methods}}

Two different techniques are deployed in this paper to study the spatial
distribution of criminal events, and their possible relationship to the
location of places of worship. Firstly, the F-function is a nearest
neighbor technique that is used to summarize the distance between a set
of points and their nearest events. This technique has found extensive
use in spatial epidemiology (see Gatrell et al., 1996) and has also been
used in spatial criminology (see, inter alia, Craglia et al., 2000;
Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001). The
second technique is the scan statistic. This method has also been
extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996)
and has found use in spatial criminology as well (see He et al., 2017a;
Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
Use of these two techniques is meant to provide insights with respect to
1) the location of violent criminal events with respect to various types
of facilities; and 2) geographically identify places of worship
associated with clusters of criminal events. These two techniques are
briefly reviewed next.

\hypertarget{quadrat-analysis}{%
\subsection{Quadrat Analysis}\label{quadrat-analysis}}

Words go here.

\hypertarget{relative-intensity}{%
\subsection{Relative Intensity}\label{relative-intensity}}

Words go here.

\hypertarget{case}{%
\section{Case Study}\label{case}}

\hypertarget{context}{%
\subsection{Context}\label{context}}

The study is of the city of Recife, the capital of the state of
Pernambuco in the Northeast region of Brazil. With a population of
1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2,
Recife is one of the main Brazilian metropolises, exerting a great
economic influence in neighboring regions. However, the city experiences
a serious problem with violent crime, and has the dubious honor of being
one of the five capital cities in Brazil with the highest homicide rates
in the period under study (Menezes et al. 2013).

In the current context, the term ``violent crime'' is an umbrella for
several forms of infractions to the penal code. Following
recommendations of the National Secretariat of Public Security of the
Ministry of Justice of 2006, these are Violent Lethal and Intentional
Crimes (VLIC; which includes intentional homicide), theft followed by
death (robbery), and corporal injury followed by death. The data on CVLI
were extracted from the Police Information System of the Secretariat of
Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most
reliable, detailed, and comprehensive information on violent deaths in
the region.

\begin{table}

\caption{\label{tab:descriptive-statistics}\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009}
\centering
\begin{tabular}[t]{lr}
\toprule
 & Percentage\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Gender of Victim}}\\
\hspace{1em}Man & 91.35\\
\hspace{1em}Woman & 8.65\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Type of Crime}}\\
\hspace{1em}Murder & 97.82\\
\hspace{1em}Robbery & 2.00\\
\hspace{1em}Body injury followed by death & 0.18\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Ethnicity of Victim}}\\
\hspace{1em}Black and White & 95.11\\
\hspace{1em}Yellow and White & 1.27\\
\hspace{1em}Not Reported & 3.62\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Age of Victim}}\\
\hspace{1em}1 to 12 years & 0.18\\
\hspace{1em}13 to 17 years & 11.31\\
\hspace{1em}18 to 30 years & 61.70\\
\hspace{1em}31 to 65 years & 25.84\\
\hspace{1em}65 years and older & 0.73\\
\hspace{1em}Not Reported & 0.24\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Weapon Used}}\\
\hspace{1em}Firearm & 87.51\\
\hspace{1em}Other & 12.49\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{data}{%
\subsection{Data}\label{data}}

The data are organized at the individual level and it is possible to
obtain information about the location, day of the week, day of the
month, period of the day, as well as gender, age, and race of the
victim. The database used in this study comprises the period from July
1, 2008 to June 30, 2010. Some descriptive statistics regarding this
dataset are reported in Table 1, where it can be seen that overall,
about 91\% of the victims of CVLI in the period analyzed were men, while
approximately 98\% of violent crimes were homicides. In addition, most
of the victims were black or brown, and the youth population between the
ages of 18 and 30 is the most affected by violent crime. Lastly, it
should be noted that about 88\% of the criminal events under analysis
involved firearms. Information about places of interest was obtained
from the National Register of Addresses for Statistical Purposes (CNEFE
- Census 2010), which lists 78,056,411 urban and rural addresses,
distributed among the 316,574 census tracts. This is the first database
of its kind produced by IBGE, and the first version was produced at the
time of the 2000 Census. The way addresses are described in the National
Register is very rich, and it is possible to identify the names of the
places of worship including their denomination. Georeferencing was used
to geolocate each place of worship. In this way, a total of 1,719 places
of worship were geolocated in the city of Recife. Figure 1 (Panel a)
shows the spatial distribution of the 1,657 CVLI crimes that occurred in
the city of Recife between 2008.2 and 2010.1. Panel b) of the figure
shows the spatial distribution of places of worship in the region.

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-1} \hfill{}

\caption{\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife}\label{fig:plot-crime}
\end{figure}

In addition to places of worship, the National Register of Addresses for
Statistical Purposes was queried to extract facilities other than places
of worship. As discussed above, the idea is to identify points of
reference that can be used as controls, having a neutral morality
profile. For the sake of the present study, we selected pharmacies, ice
cream shops, and bakeries to construct our control group. These three
types of establishments comply with the criteria of being morally
neutral and having a spatial distribution commensurate with places of
worship. Figure 2 shows the spatial distribution of the control
establishments, namely pharmacies, bakeries, and ice cream shops in the
city of Recife. Note the similarity between the maps. As expected, there
are differences in the number of points, but the locations of cases and
controls are quite similar.

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-1} \hfill{}

\caption{\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife}\label{fig:plot-events}
\end{figure}

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-1} \hfill{}

\caption{\label{fig:plot-sed}Socio-Economic Deprivation in Recife}\label{fig:plot-sed}
\end{figure}

\hypertarget{results}{%
\section{Analysis and results}\label{results}}

\hypertarget{quadrat-analysis-1}{%
\subsection{Quadrat Analysis}\label{quadrat-analysis-1}}

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-as-quintiles-1.pdf}
\caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as
Quintiles}
\end{figure}

Calculate the intensity of other ppp by quadrat:

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-quadrat-1.pdf}
\caption{\label{fig:plot-crime-quadrat}Intensity of crime by level of
Socio-Economic Deprivation; the dotted line indicates the global
intensity of crime}
\end{figure}

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-quadrat-1.pdf}
\caption{\label{fig:plot-events-quadrat}Intensity of places of worship
and commercial establishments by level of Socio-Economic Deprivation;
the dotted lines indicates the respective global intensities}
\end{figure}

\hypertarget{relative-intensity-1}{%
\subsection{Relative Intensity}\label{relative-intensity-1}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

An opposing viewpoint is offered by the theory of conservative
Protestants (Ellison, Burr, and McCall 2003). This theory argues that
religion is not always beneficial in order to guarantee social order,
and consequently to result in lower crime rates. This theory suggested
that, conservative Protestants are more tolerate with violent behavior
when associated with defense of honor, family, property, or woman.
However, high rates are not expected for all types of crimes in densely
populated areas of conservative Protestants. But what justifies this
behavior? Ellison et al.~(2003) argue that this tolerance stems from a
religious view of the world. More clearly, punishment is seen as a
legitimate form of retribution whose explanation can be attributed to
the literal interpretation of the Bible. In the specific context of
Brazil, it is important to consider that the interpretations and
meanings attributed to urban violence during the last two decades in
Brazil often had religious figures as protagonists (Birman and Machado
2012).

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

The theory of conservative Protestants, argues that conservative
Protestants are more likely to tolerate violent behavior of a defensive
or punitive character, especially when associated to defense of honor,
family, property, or woman. Our objective in the study is to explore the
question whether moral communities have a spatial dimension. To this end
we deploy an array of spatial analytical tools to investigate potential
patterns of clustering of violent crime and places of worship, using as
well other kinds of urban facilities to act as controls. The case study
is the city of Recife, in the northeast of Brazil, and analysis is based
on geographically detailed datasets of criminal events as well as
several types of urban facilities.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abdullah2018amenities}{}%
Abdullah, Snhs, F. A. Bohani, Z. A. Nazri, Y. Jeffry, M. A. Abdullah, M.
N. Junoh, and Z. A. Kasim. 2018. ``Amenities Surrounding Commercial
Serial Crime Prediction at Greater Valley and Kuala Lumpur Using K-Means
Clustering.'' Journal Article. \emph{Jurnal Teknologi} 80 (4): 43--53.
\href{\%3CGo\%20to\%20ISI\%3E://WOS:000437011000005}{\textless{}Go to ISI\textgreater{}://WOS:000437011000005}.

\leavevmode\hypertarget{ref-Becker2017analise}{}%
Becker, Kalinca Léia, and Ana Lúcia Kassouf. 2017. ``Uma análise Do
Efeito Dos Gastos Públicos Em Educação Sobre a Criminalidade No
Brasil.'' \emph{Economia E Sociedade} 26 (1): 215--42.

\leavevmode\hypertarget{ref-Birman2012violencia}{}%
Birman, Patricia, and Carly Machado. 2012. ``A Violencia Dos Justos:
Evangelicos, Midia E Periferias Da Metropole.'' \emph{Revista Brasileira
de Ciências Sociais} 27 (80): 55--69.

\leavevmode\hypertarget{ref-Brower2007spatial}{}%
Brower, A. M., and L. Carroll. 2007. ``Spatial and Temporal Aspects of
Alcohol-Related Crime in a College Town.'' Journal Article.
\emph{Journal of American College Health} 55 (5): 267--75.
\url{https://doi.org/10.3200/jach.55.5.267-276}.

\leavevmode\hypertarget{ref-Chesnut2017devoted}{}%
Chesnut, R Andrew. 2017. \emph{Devoted to Death: Santa Muerte, the
Skeleton Saint}. Oxford University Press.

\leavevmode\hypertarget{ref-Davignon2015christian}{}%
Davignon, P., and R. A. Thomson. 2015. ``Christian Colleges and
Universities as Moral Communities: The Effects of Institutional
Characteristics on Student Religiosity.'' Journal Article. \emph{Review
of Religious Research} 57 (4): 531--54.
\url{https://doi.org/10.1007/s13644-015-0214-5}.

\leavevmode\hypertarget{ref-Deryol2016crime}{}%
Deryol, R., P. Wilcox, M. Logan, and J. Wooldredge. 2016. ``Crime Places
in Context: An Illustration of the Multilevel Nature of Hot Spot
Development.'' Journal Article. \emph{Journal of Quantitative
Criminology} 32 (2): 305--25.
\url{https://doi.org/10.1007/s10940-015-9278-1}.

\leavevmode\hypertarget{ref-Donovan1986different}{}%
Donovan, Peter. 1986. ``Do Different Religions Share Moral Common
Ground?'' Journal Article. \emph{Religious Studies} 22 (3/4): 367--75.
\url{http://www.jstor.org.libaccess.lib.mcmaster.ca/stable/20006295}.

\leavevmode\hypertarget{ref-Doran2011putting}{}%
Doran, Bruce J, and Melissa B Burgess. 2011. \emph{Putting Fear of Crime
on the Map: Investigating Perceptions of Crime Using Geographic
Information Systems}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-Durrant2017religion}{}%
Durrant, Russil, and Zoe Poppelwell. 2017. ``Why Religion Matters.'' In
\emph{Religion, Crime and Punishment}, 1--17. Springer.

\leavevmode\hypertarget{ref-Eitle2011religion}{}%
Eitle, D. 2011. ``Religion and Gambling Among Young Adults in the United
States: Moral Communities and the Deterrence Hypothesis.'' Journal
Article. \emph{Journal for the Scientific Study of Religion} 50 (1):
61--81. \url{https://doi.org/10.1111/j.1468-5906.2010.01552.x}.

\leavevmode\hypertarget{ref-Ellison2003enduring}{}%
Ellison, Christopher G, Jeffrey A Burr, and Patricia L McCall. 2003.
``The Enduring Puzzle of Southern Homicide: Is Regional Religious
Culture the Missing Piece?'' \emph{Homicide Studies} 7 (4): 326--52.

\leavevmode\hypertarget{ref-Foster2010neighbourhood}{}%
Foster, S., B. Giles-Corti, and M. Knuiman. 2010. ``Neighbourhood Design
and Fear of Crime: A Social-Ecological Examination of the Correlates of
Residents' Fear in New Suburban Housing Developments.'' Journal Article.
\emph{Health \& Place} 16 (6): 1156--65.
\url{https://doi.org/10.1016/j.healthplace.2010.07.007}.

\leavevmode\hypertarget{ref-Furr2010metric}{}%
Furr-Holden, C. D. M., K. D. M. Campbell, A. J. Milam, M. J. Smart, N.
A. Ialongo, and P. J. Leaf. 2010. ``Metric Properties of the
Neighborhood Inventory for Environmental Typology (Nifety): An
Environmental Assessment Tool for Measuring Indicators of Violence,
Alcohol, Tobacco, and Other Drug Exposures.'' Journal Article.
\emph{Evaluation Review} 34 (3): 159--84.
\url{https://doi.org/10.1177/0193841x10368493}.

\leavevmode\hypertarget{ref-Groff2015informal}{}%
Groff, E. R. 2015. ``Informal Social Control and Crime Events.'' Journal
Article. \emph{Journal of Contemporary Criminal Justice} 31 (1):
90--106. \url{https://doi.org/10.1177/1043986214552619}.

\leavevmode\hypertarget{ref-He2017built}{}%
He, Li, Antonio Páez, and Desheng Liu. 2017. ``Built Environment and
Violent Crime: An Environmental Audit Approach Using Google Street
View.'' Journal Article. \emph{Computers, Environment and Urban Systems}
66: 83--95.
\url{https://doi.org/http://dx.doi.org/10.1016/j.compenvurbsys.2017.08.001}.

\leavevmode\hypertarget{ref-Hipp2007block}{}%
Hipp, J. R. 2007. ``Block, Tract, and Levels of Aggregation:
Neighborhood Structure and Crime and Disorder as a Case in Point.''
Journal Article. \emph{American Sociological Review} 72 (5): 659--80.
\url{https://doi.org/10.1177/000312240707200501}.

\leavevmode\hypertarget{ref-Hirschi1969hellfire}{}%
Hirschi, Travis, and Rodney Stark. 1969. ``Hellfire and Delinquency.''
\emph{Social Problems} 17 (2): 202--13.

\leavevmode\hypertarget{ref-Hoffmann2015religion}{}%
Hoffmann, John P. 2015. ``Religion and Crime.'' \emph{The Encyclopedia
of Crime and Punishment}, 1--5.

\leavevmode\hypertarget{ref-Johnson2011more}{}%
Johnson, Byron. 2011. \emph{More God, Less Crime: Why Faith Matters and
How It Could Matter More}. Templeton Foundation Press.

\leavevmode\hypertarget{ref-Lee2004love}{}%
Lee, M. R., and J. P. Bartkowski. 2004. ``Love Thy Neighbor? Moral
Communities, Civic Engagement, and Juvenile Homicide in Rural Areas.''
Journal Article. \emph{Social Forces} 82 (3): 1001--35.
\url{https://doi.org/10.1353/sof.2004.0044}.

\leavevmode\hypertarget{ref-Lipton2008spatial}{}%
Lipton, R., A. Banerjee, D. Levy, N. Manzanilla, and M. Cochrane. 2008.
``The Spatial Distribution of Underage Tobacco Sales in Los Angeles.''
Journal Article. \emph{Substance Use \& Misuse} 43 (11): 1597--1617.
\url{https://doi.org/10.1080/10826080802241110}.

\leavevmode\hypertarget{ref-Loukaitou2001measuring}{}%
Loukaitou-Sideris, A., R. Liggett, H. Iseki, and W. Thurlow. 2001.
``Measuring the Effects of Built Environment on Bus Stop Crime.''
Journal Article. \emph{Environment and Planning B-Planning \& Design} 28
(2): 255--80. \url{https://doi.org/10.1068/b2642r}.

\leavevmode\hypertarget{ref-Menezes2013spatial}{}%
Menezes, T., R. Silveira-Neto, C. Monteiro, and J. L. Ratton. 2013.
``Spatial Correlation Between Homicide Rates and Inequality: Evidence
from Urban Neighborhoods.'' Journal Article. \emph{Economics Letters}
120 (1): 97--99. \url{https://doi.org/10.1016/j.econlet.2013.03.040}.

\leavevmode\hypertarget{ref-Murray2013crime}{}%
Murray, J., D. R. D. Cerqueira, and T. Kahn. 2013. ``Crime and Violence
in Brazil: Systematic Review of Time Trends, Prevalence Rates and Risk
Factors.'' Journal Article. \emph{Aggression and Violent Behavior} 18
(5): 471--83. \url{https://doi.org/10.1016/j.avb.2013.07.003}.

\leavevmode\hypertarget{ref-Nie2019smoking}{}%
Nie, F. H., and X. Z. Y. Yang. 2019. ``Smoking in the Temple of the Holy
Spirit? Geographic Location Matters.'' Journal Article. \emph{Health \&
Place} 58. \url{https://doi.org/10.1016/j.healthplace.2019.05.017}.

\leavevmode\hypertarget{ref-Quick2017influence}{}%
Quick, M., J. Law, and H. Luan. 2017. ``The Influence of on-Premise and
Off-Premise Alcohol Outlets on Reported Violent Crime in the Region of
Waterloo, Ontario: Applying Bayesian Spatial Modeling to Inform Land Use
Planning and Policy.'' Journal Article. \emph{Applied Spatial Analysis
and Policy} 10 (3): 435--54.
\url{https://doi.org/10.1007/s12061-016-9191-5}.

\leavevmode\hypertarget{ref-Regnerus2003moral}{}%
Regnerus, M. D. 2003. ``Moral Communities and Adolescent Delinquency:
Religious Contexts and Community Social Control.'' Journal Article.
\emph{Sociological Quarterly} 44 (4): 523--54.
\href{\%3CGo\%20to\%20ISI\%3E://WOS:000188319800001}{\textless{}Go to ISI\textgreater{}://WOS:000188319800001}.

\leavevmode\hypertarget{ref-Reiss1951delinquency}{}%
Reiss, Albert J. 1951. ``Delinquency as the Failure of Personal and
Social Controls.'' Journal Article. \emph{American Sociological Review}
16 (2): 196--207. \url{https://doi.org/10.2307/2087693}.

\leavevmode\hypertarget{ref-Rohrbaugh1975religiosity}{}%
Rohrbaugh, John, and Richard Jessor. 1975. ``Religiosity in Youth: A
Personal Control Against Deviant Behavior.'' \emph{Journal of
Personality} 43 (1): 136--55.

\leavevmode\hypertarget{ref-Stark1996religion}{}%
Stark, R. 1996. ``Religion as Context: Hellfire and Delinquency One More
Time.'' Journal Article. \emph{Sociology of Religion} 57 (2): 163--73.
\url{https://doi.org/10.2307/3711948}.

\leavevmode\hypertarget{ref-Stroope2018moral}{}%
Stroope, S., and J. O. Baker. 2018. ``Whose Moral Community?
Religiosity, Secularity, and Self-Rated Health Across Communal Religious
Contexts.'' Journal Article. \emph{Journal of Health and Social
Behavior} 59 (2): 185--99.
\url{https://doi.org/10.1177/0022146518755698}.

\leavevmode\hypertarget{ref-Topalli2013god}{}%
Topalli, V., T. Brezina, and M. Bernhardt. 2013. ``With God on My Side:
The Paradoxical Relationship Between Religious Belief and Criminality
Among Hardcore Street Offenders.'' Journal Article. \emph{Theoretical
Criminology} 17 (1): 49--69.
\url{https://doi.org/10.1177/1362480612463114}.

\leavevmode\hypertarget{ref-Traunmuller2011moral}{}%
Traunmuller, R. 2011. ``Moral Communities? Religion as a Source of
Social Trust in a Multilevel Analysis of 97 German Regions.'' Journal
Article. \emph{European Sociological Review} 27 (3): 346--63.
\url{https://doi.org/10.1093/esr/jcq011}.

\leavevmode\hypertarget{ref-Unodc2019executive}{}%
United Nations Office on Drugs and Crime. 2019a. ``Global Study on
Homicide 2019: Executive Summary.'' UNODC Vienna.

\leavevmode\hypertarget{ref-Unodc2019development}{}%
---------. 2019b. ``Global Study on Homicide 2019: Homicide, Development
and the Sustainable Development Goals.'' UNODC Vienna.

\leavevmode\hypertarget{ref-Warner2019neighborhood}{}%
Warner, B. D., and R. H. Konkel. 2019. ``Neighborhood Churches and Their
Relationship to Neighborhood Processes Important for Crime Prevention.''
Journal Article. \emph{Journal of Urban Affairs}.
\url{https://doi.org/10.1080/07352166.2019.1581030}.

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}