### 2019-10-14 Moral_Communities_and_Crime_v1.Rmd
Completed first draft of methods and started working on the sections on context/data and analysis and results. Changed the name of the RData file to Moral_Communities_Crime_Recife.RData

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Aye B. Cee
  address: A Department, University of XXX
  email: abc@example.edu

- name: Dee E. Eff
  address: A Department, University of XXX
  email: def@example.edu

- name: Gee H. Eye
  address: A Department, University of XXX
  email: ghi@example.edu

- name: Jay K. Elle
  address: Some School, University of WWW
  email: jkl@school.edu

keywords:
- Moral communities
- Crime
- Places of worship
- Point pattern analysis
- Intensity

abstract: |
  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries. With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities. Moral communities, a concept closely related to informal social controls, are of interest in countries and regions where formal controls are weak and ineffective. The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship. Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns. Analysis is condcuted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets). The results suggest that Catholic and Spiritist places of worship do not project moral communities geographically more than, say, ice cream shops. The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher than for any of our other referential events.

bibliography: bibliography.bib
output: rticles::springer_article
---

<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following file is needed: `Moral_Communities_and_Crime.RData`. For reference: this data file was created using R notebook `00. Data Preprocessing.Rmd`. The initial analysis was done in R notebook `01. Analysis.Rmd`. Neither of these two files are needed for this document. -->

```{r load-packages, include=FALSE}

# Load packages used in this document

library(tidyverse)
library(kableExtra)
library(sf)
library(spatstat)
library(maptools)
library(ggthemes)
library(raster)
library(fasterize)
```

```{r load-data, include=FALSE}

# To run this notebook, a file `Moral_Communities_and_Crime.RData` is needed. This file contains the following objects:
# 
# - events_recife.sf: a simple features object with points for violent crimes, churches (Catholic, Evangelical, Spiritist) and commercial establishments (ice cream shops, pharmacies, supermarkets) in the city of Recife.
# 
# - recife.sf: a simple features object with the boundaries of the city of Recife
# 
# - setores.sf: a simple features object with setores (census tracts) and selected socio-economic and demographic information for the city of Recife, including:
#   - Population density in persons per sq.km
#   - Median income in setor
#   - Proportion of population in setor who are unemployed
#   - Proportion of households with a single mother
#   - Proportion of households with a young single mother, a woman between the ages of 15 and 25
#   - Proportion of population in poverty
# 
# Read the RData file:

load("Moral_Communities_and_Crime_Recife.RData")
```

```{r data-preparation, include=FALSE, cache=TRUE}

# In preparation for doing point pattern analysis, convert the `recife.sf` object to an `owin` object for use in `spatstat`:

recife.owin <- recife.sf %>% as("Spatial") %>% as("owin")
plot(recife.owin)

# Convert the events to a `ppp` object:

events_recife.ppp <- events_recife.sf %>% 
  st_coordinates() %>% 
  data.frame() %>%
  mutate(Type = events_recife.sf$Type) %>%
  as.ppp(W = recife.owin)

#Split the point pattern into each type of event: crime, Catholic place of worship, Evangelical place of worship, Spiritist place of worship, ice cream shop, pharmacy, and supermarket

events_recife_list <- split(events_recife.ppp)
```

```{r socio-economic-deprivation, include=FALSE, cache=TRUE}

# Here an indicator of socio-economic deprivation is obtained using principal component analysis

# PCA is implemented in R as `prcomp`:
setores.pca <- setores.sf %>% 
  st_drop_geometry() %>%
  dplyr::select(median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
  prcomp(center = TRUE, scale = TRUE)

# The loadings of the pca are as follows:
setores.pca$rotation

# Socio-economic deprivation is a combination of high poverty, high proportion of single mothers, especially young women between the ages of 15-25, high proportion of unemployed people, and low median income.

# Various diagnostics of the results are possible, such as the variance:

std_dev <- setores.pca$sdev

# Compute variance

pr_var <- std_dev^2

# Check variance of the components

pr_var[1:5]

# And the proportion of the variance:

pr_var/sum(pr_var)

# Once that there is principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:

setores.sf$SED <- setores.pca$x[,1]

# Normalize the SDE factor (with the addition of a small correction; the `rhohat` function does not like values <= 0):

setores.sf <- setores.sf %>% mutate(SED = (SED - min(SED) + 0.1)/(max(SED) + 0.1 - min(SED)))
```

# Introduction {#intro}

Violent crime is a widespread phenomenon with negative impacts on many spheres of life and society. Although the rate of homicides worldwide has grown at a slower rate than the population in the past few decades, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive]. Furthermore, variations in the prevalence of violent crime tend to be extremely uneven. The Americas, with a population of approximately 793.8 million people in 2019 (or about 10.3% of the world population), accounted for approximately 173,000 homicides in 2017, or approximately 37.3% of all homicides in the world. Within the Americas, Brazil (along with Venezuela, Colombia, and Mexico) is one of the largest countries in the region with high homicide rates. And there are notable variations at the sub-national and intra-metropolitan levels as well. In fact, understanding the variations in the prevalence of violent crime is recognized as key to achieving policy goals:

>> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so
achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p. 35]

This concern is not iddle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals. Violent crime, furthermore, tends to be more accute in those regions most in need of development. On the one hand, violent crime represents an economic and social drag [@Becker2017analise]. The impacts are multifaceted, since they affect people's well-being, either through the loss of human life, mental health, and limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development]. In turn, these negative effects combine to create the unfortunate conditions that tend to breed criminal behavior, thus creating a vicious cycle of economic disadvantage and violence [@Unodc2019development]. It is not surprising, given these pecularities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].

Added to the scenario, formal social control institutions in countries like Brazil leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up reducing the deterrent power of the justice system. Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial]. This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborbood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].

In order to understand the factors that can potentially help to deter crime, beyond formal control, it is important to identify empirical regularities. Criminological factors include concrete elements (such as te presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the soci costs of deviant behavior as well as family supervision. Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets [e.g., @Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.

With the above considerations in mind, the objective of this paper is to investigate whether and how places of workship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension. The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast. Recife is a large and important metropolitan area in a historically poor region, and afflicted by high levels of violent crime, having the dubious honor of being one of five state capitals with the highest rates of homicide in the period under study [@Menezes2013spatial]. The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments that serve as controls, on the other. Dissagregated data allows us to analyze the relevant point patterns at a very high level of resolution.

After this introductory section, the rest of this paper is structured as follows: Section \ref{background} reviews some relevant theoretical perspectives on religion and crime; then Section \ref{methods} presents the empirical strategy used in the work. Section \ref{case} presents the context as well as the data used in the study; Section \ref{results} presents the analysis, whereas the results are discussed in Section \ref{discussion}. Finally, some concluding remarks are offered in Section \ref{conclusions}.

# Background: Theoretical Perspectives on Religion and Crime {#background}

Why do humans behave morally? For millennia, it has been the role of religion to provide the basic tenets of morally acceptable behavior: thou shalt not kill et al., and their equivalent in many world religions [e.g., @Donovan1986different]. Enforcement of such tenets implies different mechanisms, including _sin_, _haram_, _karma_, and _tapu_, with punishment delivered by hellfire and exile to _Jahannam_ or _Gehenna_, to mention just some choice places of torment. In addition to acting as _de facto_ social policy for much of history, many of these religious tenets still have the force of law in many cultures and regions - and even where they do not, they are held by some researchers and policy experts to be helpful complements to reduce crime in any case [e.g., @Durrant2017religion; @Johnson2011more]. The hypothesis that religion can act as a factor that deters and reduces criminal behavior, therefore, has prompted the scientific study of the effectiveness of religion on moral behavior [@Hoffmann2015religion].

An early effort to theorize the effect of religion on criminal behavior was the _hellfire hypothesis_ of Hirschi and Stark [-@Hirschi1969hellfire]. The focus of this hypothesis is the threat of extratemporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager). Accordingly, Hirsch and Stark [-@Hirschi1969hellfire] posit that negative correlations between crime and religion are a consequence of a sense of commitment with normative values, a commitment that is ritually reinforced in a regular fashion (e.g., the liturgical rite of peace in Catholic Mass). Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity], for instance, note that by attributing to divinity (or some other supra-terrenal entity) the supreme force of punishment, religion helps to build a sense of obduracy, or "hardening" against temptation.

Despite its intuitive appeal, research has not provided much direct evidence for the hypothesis of hellfire [@Hoffmann2015religion, p. 1]. In a bizarre twist, even mainstream doctrines (such as absolution) may in fact encourage beliefs that neutralize the fear of _terrenal_ punishment, and thus turn out to be criminogenic [e.g., @Topalli2013god]. More deviant cases can even recast criminal behavior as a form of spiritual insurgency [@Chesnut2017devoted]. Counterexamples like this notwithstanding, researchers remain open to the possibility that "religion may still serve as a social control mechanism by encouraging conventional beliefs, monitoring behaviors, enhancing family attachments, or providing conventional activities" [@Hoffmann2015religion, p. 1]. In this way, the hypothesis of _moral communities_ [@Stark1996religion] recognizes that social integration is essential for increasing social control, thus reducing the practice of behavior that is not in accordance with current norms. Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective” [@Reiss1951delinquency, p. 196; cited in @Groff2015informal, p. 91].

The hypothesis of moral communities has over the years been used to examine a variety of outcomes of interest. Recent examples include Stroope and Baker's [-@Stroope2018moral] exploration of religiosity and self-rated health; Davignon and Thomson [-@Davignon2015christian] with their research on institutional context and the religiosity of students; and the study of religion as a source of trust of Traunmuller [-@Traunmuller2011moral]. This is in addition, of course, to numerous studies on criminal behavior such as Eitle [-@Eitle2011religion], an author who explored the deterrence power of religion on gambling; Lee and Bartkowski's [-@Lee2004love] investigation of juvenile homicide in rural areas; and the research of Regnerus [-@Regnerus2003moral] on adolescent delinquency. The hypothesis of moral communities as a form of informal social control, on the other hand, has been less studied from a geographical perspective, and it is only recently that has attracted the attention of researchers. Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plaussibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county. In this way, Nie and Yang [-@Nie2019smoking] remark on a recent paper on the lack of research conducted to study how the religious context of a geographical area (e.g., a county) may influence (youth smoking) behavior (p. 2). This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory are thought to occur at relatively small geographical scales.

The preceding review, as well as other research [-@Malleson2019identifying], makes clear the need for research at various geographical scales. From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood. In what follows, we propose to adopt a much more disaggregated approach by considering the intensity of criminal events with respect to places of worship. This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.

# Empirical Strategy {#methods}

As noted above, previous research that has investigated the presence of churches from a geographical perspective has used data aggregated at different scales. In the case of Nie and Yang [-@Nie2019smoking], the unit of analysis was the county, whereas Warner and Konkel [-@Warner2019neighborhood] used Census Block Groups, a much smaller unit of analysis. Research in spatial criminology, on the other hand, has studied criminal events as point patterns since at least the work of Levine et al. [-@Levine1986crime] investigated the concentration of criminal events in the proximity of bus stops. Since then, many other works have applied tools of spatial point pattern analysis to investigate the empirical properties of the distribution of crime. This includes Craglia et al. [-@Craglia2000comparative] who used the Ripley's K-function [@Ripley1976second] to investigate clustering processes, and Rogerson and Sun [-@Rogerson2001spatial] who applied nearest neighbor techniques to the study of arson. More recent studies of crime as point patterns are Nakaya and Yano [-@Nakaya2010visualizing], Kiani et al. [-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio].

Readers interested in point pattern analysis techniques are urged to consult the still valuable Bailey and Gatrell [-@Bailey1995interactive] or for an up-to-date and in-depth coverage of the topic, the excellent text by Baddeley et al. [-@Baddeley2015spatial]. In this paper we will concentrate on one property of point processes, namely the intensity. Suppose that the outcome of interest is the number of events per unit area $Y(\mathcal{A})$. This could be, for instance, the number of criminal events observed in an arbitrary area. The intensity of a point pattern then is as follows:
$$
\lambda(s)=\lim_{\text{d}s \to 0}\frac{E[Y(\text{d}s)]}{\text{d}s}
$$

where $\lambda(s)$ is the intensity of the process at point $s$, given by the expected number of events in a small area $\text{d}s$ around $s$, as the area becomes arbitrarily small. If the process is homogeneous (i.e., the intensity is a constant over space), an apt estimator of the intensity is the global intensity, simply the number of events divided by the area of the region under analysis:
$$
\hat{\lambda} = \frac{n}{|\mathcal{A}|}
$$

When the process is not homogeneous, other estimators are more appropriate. In the analysis to follow we use two estimators of intensity: conditional quadrat counts and relative distribution estimate. These techniques are briefly discussed next.

## Conditional Quadrat Counts

Quadrat counts is a relatively simple technique to analyze spatial variations in the intensity of a point process. It operates by partitioning the region under study $\mathcal{A}$ into subregions $\mathcal{A}_1,\cdots,\mathcal{A}_m$. These subregions are mutually exclusive, and their union is identical to $\mathcal{A}$. In the simplest case, the subregions have equal area. The number of events within each subregion (i.e., $n_{\mathcal{A}}$) divided by the area of the subregion (i.e., $|\mathcal{A}|$) is a local estimate of the intensity:
$$
\lambda(\mathcal{A}_i)=\frac{n_{\mathcal{A}_i}}{|\mathcal{A}_i|}
$$

A test for homogeneity consists of assessing whether the intensity of the point process at the quadrats is uniform:
$$
\lambda(\mathcal{A}_1)=\cdots=\lambda(\mathcal{A}_m)
$$

Using quadrats of equal size is a convenient simplication, but in principle the areas could be different - in which case the count of events would be proportional to the area of the quadrat under homogeneity. An interesting variation of this technique is conditional quadrat counts, whereby the partition of regions $\mathcal{A}_1,\cdots,\mathcal{A}_m$ is done to reflect an underlying covariate of interest, say $Z$. By introducing a covariate as a partitioning criterion, it becomes possible to calculate estimators of the intensity of quadrats at different levels of the value of the covariate. For exploratory purposes, we can plot average intensity by quadrat, and compare to the global intensity.

## Relative Distribution Estimate

Conditional quadrat counts allow us to explore whether the intensity of the process depends on a covariate $Z$. A different way of expressing this is as follows:
$$
\lambda(s)=\rho\big(Z(s)\big)
$$

In this case, the intensity is a function $\rho$ that maps how the intensity depends on covariate $Z$. Non-parametric estimation of $\rho$ uses the ratio of the density of covariate values at the locations of the point process, relative to the _spatial distribution function_ $G$, the density of covariate values at random locations [see @Baddeley2015spatial, p. 179].

The density of covariate values at the locations of the is obtained by means of a kernel density estimator, for example:
$$
\lambda_0(s) = \frac{1}{|\mathcal{A}|}\sum_{i = 1}^n\kappa(Z(s_i) - z)
$$
where $s_i$ are the locations of the points.

On the other hand, $G$ which is the cumulative distribution function of $Z$ at random point $S$ uniformly distributed in $\mathcal{A}$  :
$$
G(z) = \frac{1}{|\mathcal{A}|} \int_{\mathcal{A}}\text{I}\{Z(s)\leq z\}\text{d}s
$$
where $\text{I}$ is an indicator function that takes the value of $1$ if the argument is true, and the value of $0$ otherwise. In practice, the spatial distribution function is approximated based on a discretization of space using a fine grid of pixels as follows:
$$
G(z) = \frac{\#\{\text{pixels }s:Z(s)\leq z\}}{\# \text{pixels}}
$$

Therefore, an estimator of $\rho$ is as follows:
$$
\rho(z) = \frac{1}{|\mathcal{A}|G'(z)}\sum_{i=1}^n\kappa(Z(s_i) - z)
$$
where the derivative $G'$ can be approximated by differentiating a smoothed estimated of $G$. Other estimators are discussed by Baddeley et al. [@Baddeley2015spatial, p. 180].

It is possible to adjust the relative distribution estimate by means of a baseline; a baseline in this case can be a function of other covariates that might confound the estimates of the relative distribution, so that the relative intensity $\lambda(s)/B(s)$ can be assumed to depend only on covariate $Z$. Therefore:
$$
\lambda(s) = \rho(Z(s))B(s)
$$
where $B(s)$ is a baseline function. It can be seen that $\rho(z)=1$ is the baseline intensity; values of $\rho(z)>1$ correspond to intensities higher than the baseline _as a function of_ $Z(u)$, whereas values of $\rho(z)<1$ correspond to intensities lower than the baseline, again as a function of $Z(u)$.

# Case Study: Context and Data {#case}

The study is of the city of Recife, the capital of the state of Pernambuco in the Northeast region of Brazil. With a population of 1,550,390 million inhabitants in 2010, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions. However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].

In the current context, the term "violent crime" is an umbrella for several forms of infractions to the penal code. Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (VLIC; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death. The data on LIVC were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.

```{r table-descriptive-statistics, echo=FALSE}
descriptive_stats <- data.frame(Attribute = c("Man",
                                              "Woman",
                                              "Murder",
                                              "Robbery",
                                              "Body injury followed by death",
                                              "Black and White",
                                              "Yellow and White",
                                              "Not Reported",
                                              "1 to 12 years",
                                              "13 to 17 years",
                                              "18 to 30 years",
                                              "31 to 65 years",
                                              "65 years and older",
                                              "Not Reported",
                                              "Firearm",
                                              "Other"),
                                Percentage = c(91.35,
                                               8.65,
                                               97.82,
                                               2.00,
                                               0.18,
                                               95.11,
                                               1.27,
                                               3.62,
                                               0.18,
                                               11.31,
                                               61.70,
                                               25.84,
                                               0.73,
                                               0.24,
                                               87.51,
                                               12.49))

kable(descriptive_stats, 
      "latex", 
      booktabs = TRUE,
      digits = 3,
      caption = "\\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009",
      col.names = c("", "Percentage")) %>%
  kable_styling() %>%
  pack_rows("Gender of Victim", 1, 2) %>%
  pack_rows("Type of Crime", 3, 5) %>%
  pack_rows("Ethnicity of Victim", 6, 8) %>%
  pack_rows("Age of Victim", 9, 14) %>%
  pack_rows("Weapon Used", 15, 16)
```

The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim. The database used in this study comprises the period from July 1, 2008 to June 30, 2010. Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men, while approximately 98% of violent crimes were homicides. In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime. Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms. Figure \ref{fig:plot-crime} shows the spatial distribution of the 1,657 LIVC crimes that occurred in the city of Recife between July, 2008 and June, 2010.

```{r fig-plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Location of Lethal and Intentional Violent Crime in Recife, July, 2008 - June, 2009"}

#We can experiment with plotting the events:

ggplot(data = events_recife.sf %>% filter(Type == "Violent Crime")) + 
  geom_sf(color = "red") +
  geom_sf(data = recife.sf, fill = NA) + 
  theme_tufte()
```

Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts. This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census. The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination. Georeferencing was used to geolocate each place of worship. In this way, a total of 1,719 places of worship were geolocated in the city of Recife.

In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship. As discussed above, the idea is to identify points of reference that can be used as controls, having a neutral morality profile. For the sake of the present study, we selected pharmacies, ice cream shops, and bakeries to construct our control group. These three types of establishments comply with the criteria of being morally neutral and having a spatial distribution commensurate with places of worship.
Figure 2 shows the spatial distribution of the control establishments, namely pharmacies, bakeries, and ice cream shops in the city of Recife. Note the similarity between the maps. As expected, there are differences in the number of points, but the locations of cases and controls are quite similar.

```{r fig-plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Location of Places of Worship and Commercial Establishments in Recife"}

#We can experiment with plotting the events:

ggplot(data = events_recife.sf %>% filter(Type != "Violent Crime")) + 
  geom_sf(aes(color = Type, fill = Type)) +
  geom_sf(data = recife.sf, fill = NA) + 
  facet_wrap(~ Type, nrow = 2) +
  scale_fill_viridis_d(option = "viridis", direction = 1) +
  scale_color_viridis_d(option = "viridis", direction = 1) +
  theme_tufte()  +
  theme(axis.text.x = element_text(angle = 90))

```

```{r summary-point-patterns, include=FALSE}

# Extract the intensity of each point pattern for the study area

summary_crime <- summary(events_recife_list$`Violent Crime`)
summary_catholic <- summary(events_recife_list$Catholic)
summary_evangelic <- summary(events_recife_list$Evangelical)
summary_spiritist <- summary(events_recife_list$Spiritist)
summary_ice_cream <- summary(events_recife_list$`Ice Cream Shop`)
summary_pharmacy <- summary(events_recife_list$Pharmacy)
summary_supermarket <- summary(events_recife_list$Supermarket)
```

# Analysis and results {#results}

The analysis is implemented using the `R` statistical computing language, and documents to replicate the analysis, in addition to all data necessary, can be obtained from this anonymous shared folder:

https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_

## Socio-Economic Deprivation

```{r sed-descriptive-statistics, echo=FALSE}

# Create a dataframe with the descriptive statistics of the socio-economic and demographic variables used for the creation of the Socio-Economic Deprivation index. The descriptive statistics are the min, 2nd quartile, median, mean, 3rd quartile, max, and include as well the loadings of the first principal component for discussion

sed_descriptives <- data.frame(variable = c("Median Income", 
                    "Proportion Unemployed", 
                    "Proportion Poverty", 
                    "Proportion Single Mother", 
                    "Proportion Young Single Mother"),
           min = c(min(setores.sf$median_income),
                   min(setores.sf$prop_unemployed),
                   min(setores.sf$prop_poverty),
                   min(setores.sf$prop_single_mother),
                   min(setores.sf$prop_young_single_mother)),
           second_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
                        quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[2],
                        quantile(setores.sf$prop_poverty, c(0, 0.25, 0.5, 0.75, 1))[2],
                        quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2],
                        quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2]),
           median = c(median(setores.sf$median_income),
                      median(setores.sf$prop_unemployed),
                      median(setores.sf$prop_poverty),
                      median(setores.sf$prop_single_mother),
                      median(setores.sf$prop_young_single_mother)),
           mean = c(mean(setores.sf$median_income),
                    mean(setores.sf$prop_unemployed),
                    mean(setores.sf$prop_poverty),
                    mean(setores.sf$prop_single_mother),
                    mean(setores.sf$prop_young_single_mother)),
           third_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
                       quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[4],
                       quantile(setores.sf$prop_poverty, c(0, 0.25, 0.5, 0.75, 1))[4],
                       quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4],
                       quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4]),
           max = c(max(setores.sf$median_income),
                   max(setores.sf$prop_unemployed),
                   max(setores.sf$prop_poverty),
                   max(setores.sf$prop_single_mother),
                   max(setores.sf$prop_young_single_mother)),
           pca_loadings = as.numeric(setores.pca$rotation[c(1, 2, 5, 3, 4), 1]))
```

```{r fig-plot-sed, echo=FALSE, include = FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
ggplot(data = setores.sf) + 
  geom_sf(aes(fill = SED),
          color = NA) +
  scale_fill_viridis_c(option = "plasma", direction = 1, values = c(0, 0.05, 0.2, 0.3, 0.4, 1)) +
  theme_tufte()
```

The first step in our analysis is to obtain an indicator of Socio-Economic Deprivation (SED). Socio-economic deprivation is known to correlated positively with crime, as it often reflects relevant criminogenic factors such as poverty and family disruption [@He2015temporal]. In the present case, we have for the city of Recife information about the variables shown in Table \ref{tab:sed-descriptive-statistics} at the level of _setores_, a small Brazilian census geography.

As seen in the table, Recife is a city with large socio-economic and demographic disparities; for example, the _setor_ with the highest median income has a median income that is `r prettyNum(filter(sed_descriptives, variable == "Median Income")["max"]/filter(sed_descriptives, variable == "Median Income")["min"] * 100, digits = 2, big.mark = ",")`% higher than the median income in the _setor_ with the lowest median income. Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed. Likewise, there are _setores_ where almost 70% of the population lives in poverty. In addition to these economic indicators, two variables are used to represent family disruption, the proportion of families whose head is a single mother, and the proportion of families whose head is a _young_ single mother, that is, a woman between the ages of 15 and 25. As can be seen, there are _setores_ with approximately 70% of households led by single mothers, and of these, over 21% are led by younger women, indicating a high degree of family disruption.

For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation. The loadings of the first factor, which accounts for almost 54% of the variance, are shown in Table \ref{tab:sed-descriptive-statistics}. The loadings of this factor indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.

```{r table-sed-descriptive-statistics, echo=FALSE}

# Render table with descriptives of socio-economic and demographic variables

kable(sed_descriptives,
      "latex",
      booktabs = TRUE,
      digits = 2,
      escape = FALSE,
      caption = "\\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife",
      col.names = linebreak(c("Variable", 
                              "Min", 
                              "2nd Quartile", 
                              "Median", 
                              "Mean", 
                              "3rd Quartile", 
                              "Max", "PC Factor 1\n Loadings")),
      row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
  footnote(general = c("The variable for young single mothers is for women aged 15-25",
                       "The first principal component accounts for 53.91% of the variance"))

```

Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator. The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases even completely surrounded, by low deprivation areas. This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].

```{r rasterize-sed, include=FALSE}

# To use the SED indicator for conditional quadrat counts and relative distribution, we need to rasterize and convert this variable to an image:

# First, we need a base raster; this will be the boundaries of the study area. After some trial and error with different values of number of rows and columns, these values create pixels that are as close to squares as possible, while also being relatively small

recife.r <- raster(recife.sf, nrows = 400, ncols = 276)

# The raster `recife.r` is the template for rasterizing other items, in this case socio-economic deprivation. Use package `fasterize` to rasterize an `sf` object.

sed.r <- fasterize(setores.sf, recife.r, field = "SED")

# Once that the object is a raster, it can be converted to a `spatstat` `img` class object:
sed.img <- as.im(sed.r)
```

```{r quadrat-analysis-preliminaries, include=FALSE}

# To do quadrat analysis by different levels of SED we need the quintiles of this variable (other criteria could be used to slice the distribution, e.g., standard deviations, but this one is intuitive and simple)

sed5 <- quantile(sed.img, c(0, 0.2, 0.4, 0.6, 0.8, 1))

# Next we can cut the `sed.im` using the quintiles:

sed_cut5.img <- cut(sed.img, breaks = sed5, include.lowest = TRUE, labels = c(1:5))
```

```{r tesselate-sed5, include=FALSE}

# To do quadrat analysis we need to tesellate the cut image:

V <- tess(image = sed_cut5.img)

# This object is used for the analysis, but it is not very neat for plotting purposes. Lets convert instead to an `sf` object. First, extract and expand the coordinates:

coords_V <- expand.grid(X = V$image$xcol, Y = V$image$yrow)

# Next, extract the values of the cells and convert to a vector:

sed5_V <- as.vector(t(V$image$v))

# Join the coordinates with the values in a dataframe and drop all NAs, before converting to `sf` and adding coordinate referencing system information (crs) to ensure that its projection information is there and compatible with the rest of `sf` objects (i.e., recife.sf, setores.sf, events_recife.sf):

sed5.sf <- data.frame(coords_V, sed5_V) %>% 
  drop_na() %>% transmute(X, Y, SED = factor(sed5_V, 
                                              levels = c(1:5),
                                              labels = c("Low", 
                                                         "Medium Low", 
                                                         "Medium", 
                                                         "Medium High", 
                                                         "High"))) %>%
  st_as_sf(coords = c("X", "Y")) %>%
  st_set_crs(value = st_crs(recife.sf))
```

```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation in Recife classified by quintiles" }
ggplot(data = sed5.sf) + 
  geom_sf(aes(fill = SED, color = SED)) +
  geom_sf(data = recife.sf, fill = NA) +
  scale_fill_viridis_d(option = "plasma", direction = 1) +
  scale_color_viridis_d(option = "plasma", direction = 1) +
  guides(fill = guide_legend(reverse = TRUE),
         color = guide_legend(reverse = TRUE)) +
  theme_tufte()
```

## Conditional Quadrat Analysis

Calculate the intensity of other ppp by quadrat:
```{r quadrat-analysis, include=FALSE}

# Calculate the intensity of the various point patterns by quadrat:

# Crime

crime5 <- quadratcount(events_recife_list$`Violent Crime`, tess = V)
crime5_intensity <- intensity(crime5)

# Catholic

catholic5 <- quadratcount(events_recife_list$Catholic, tess = V)
catholic5_intensity <- intensity(catholic5)

# Evangelical

evangelic5 <- quadratcount(events_recife_list$Evangelical, tess = V)
evangelic5_intensity <- intensity(evangelic5)

# Spiritist

spiritist5 <- quadratcount(events_recife_list$Spiritist, tess = V)
spiritist5_intensity <- intensity(spiritist5)

# Ice cream

ice_cream5 <- quadratcount(events_recife_list$`Ice Cream Shop`, tess = V)
ice_cream5_intensity <- intensity(ice_cream5)

# Pharmacies

pharmacies5 <- quadratcount(events_recife_list$Pharmacy, tess = V)
pharmacies5_intensity <- intensity(pharmacies5)

# Supermarkets

super_markets5 <- quadratcount(events_recife_list$Supermarket, tess = V)
supermarkets5_intensity <- intensity(super_markets5)
```

```{r instensity-dataframe, include=FALSE}

# Collect in a data.frame for convenience of plotting:

intensity_all <- data.frame(SED = rep(c("Low", 
                                        "Middle Low", 
                                        "Middle", 
                                        "Middle High", 
                                        "High"), 
                                      times = 7),
                            Type = rep(c("Violent Crime", 
                                         "Catholic",
                                         "Evangelical",
                                         "Spiritist",
                                         "Ice Cream",
                                         "Pharmacies",
                                         "Super Markets"), 
                                       each = 5),
                            Intensity = c(crime5_intensity,
                                          catholic5_intensity,
                                          evangelic5_intensity,
                                          spiritist5_intensity,
                                          ice_cream5_intensity,
                                          pharmacies5_intensity,
                                          supermarkets5_intensity),
                            Global_Intensity = rep(c(summary_crime$intensity,
                                                     summary_catholic$intensity,
                                                     summary_evangelic$intensity,
                                                     summary_spiritist$intensity,
                                                     summary_ice_cream$intensity,
                                                     summary_pharmacy$intensity,
                                                     summary_supermarket$intensity),
                                                   each = 5)) %>%
  mutate(SED = factor(SED, levels = c("Low", 
                                        "Middle Low", 
                                        "Middle", 
                                        "Middle High", 
                                        "High"),
                      ordered = TRUE),
         Type = factor(Type, levels = c("Violent Crime", 
                                        "Catholic",
                                        "Evangelical",
                                        "Spiritist",
                                        "Ice Cream",
                                        "Pharmacies",
                                        "Super Markets")))
```

```{r plot-crime-quadrat, echo=FALSE, fig.width=3, fig.asp=1 , cache = TRUE, fig.cap="\\label{fig:plot-crime-quadrat}Intensity of crime by level of Socio-Economic Deprivation; the dotted line indicates the global intensity of crime"}

#Plot the intensity of crime:

ggplot(data = intensity_all %>% filter(Type == "Violent Crime"), 
       aes(x = SED, y = Intensity, fill = Type)) + 
  geom_col(fill = "red") +
  xlab("Socio-Economic Deprivation") +
  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
  #facet_wrap(Type ~.) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r plot-events-quadrat, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-events-quadrat}Intensity of places of worship and commercial establishments by level of Socio-Economic Deprivation; the dotted lines indicates the respective global intensities"}

#Plot the intensity of point patterns:

ggplot(data = intensity_all %>% filter(Type != "Violent Crime"), 
       aes(x = SED, y = Intensity, fill = Type)) + 
  geom_col() +
  xlab("Socio-Economic Deprivation") +
  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
  scale_fill_viridis_d(option = "viridis", direction = 1) +
  scale_color_viridis_d(option = "viridis", direction = 1) +
  facet_wrap(Type ~.) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

## Relative Distribution

```{r distance-functions, include=FALSE, cache = TRUE}

# Estimation of the relative intensity (see Baddeley et al. (2015) Spatial Point Pattern, pp. 179-183).

# This needs that we convert the point patterns for places of worship to distance functions. This way, the covariates for the relative distribution estimate are distance estimates to places of worship. The function `distfun` is used for this:

z_catholic <- distfun(events_recife_list$Catholic)
z_evangelical <- distfun(events_recife_list$Evangelical)
z_spiritist <- distfun(events_recife_list$Spiritist)
z_ice_cream <- distfun(events_recife_list$`Ice Cream Shop`)
z_pharmacies <- distfun(events_recife_list$Pharmacy)
z_super_markets <- distfun(events_recife_list$Supermarket)
```

## Relative Distribution with a Baseline Function

# Discussion {#discussion}

An opposing viewpoint is offered by the theory of conservative Protestants [@Ellison2003enduring]. This theory argues that religion is not always beneficial in order to guarantee social order, and consequently to result in lower crime rates. This theory suggested that, conservative Protestants are more tolerate with violent behavior when associated with defense of honor, family, property, or woman. However, high rates are not expected for all types of crimes in densely populated areas of conservative Protestants. But what justifies this behavior? Ellison et al. [-@Ellison2003enduring] argue that this tolerance stems from a religious view of the world. More clearly, punishment is seen as a legitimate form of retribution whose explanation can be attributed to the literal interpretation of the Bible. In the specific context of Brazil, it is important to consider that the interpretations and meanings attributed to urban violence during the last two decades in Brazil often had religious figures as protagonists [@Birman2012violencia].

# Conclusions {#conclusions}

The theory of conservative Protestants, argues that conservative Protestants are more likely to tolerate violent behavior of a defensive or punitive character, especially when associated to defense of honor, family, property, or woman. Our objective in the study is to explore the question whether moral communities have a spatial dimension. To this end we deploy an array of spatial analytical tools to investigate potential patterns of clustering of violent crime and places of worship, using as well other kinds of urban facilities to act as controls. The case study is the city of Recife, in the northeast of Brazil, and analysis is based on geographically detailed datasets of criminal events as well as several types of urban facilities.

# References