### 2019-10-15 Moral_Communities_and_Crime_Title_Page_v1.Rmd
First complete draft of paper and title page

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Edivaldo Constantino das Neves Júnior 
  address: A Department, University of XXX
  email: edivaldoconstantino@gmail.com

- name: Breno Caldas de Araujo
  address: A Department, University of XXX
  email: def@example.edu

- name: Tatiane Almeida de Menezes
  address: Programa de Pós-Graduação em Gestão e Economia da Saúde, Universidade Federal de Pernambuco
  email: tatianedemenezes@gmail.com

- name: Antonio Paez
  address: School of Geography and Earth Sciences, McMaster University
  email: paezha@mcmaster.ca

keywords:
- Moral communities
- Crime
- Places of worship
- Point pattern analysis
- Intensity

abstract: |
  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries. With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities. Moral communities, a concept closely related to informal social controls, are of interest in countries and regions where formal controls are weak and ineffective. The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship. Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns. Analysis is condcuted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets). The results suggest that Catholic and Spiritist places of worship do not project moral communities geographically more than, say, ice cream shops. The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher than for any of our other referential events.

bibliography: bibliography.bib
output: rticles::springer_article
---