### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime_Title_Page.Rmd
Revisions before submission to Journal of Quantitative Criminology

---
title: Spatial clustering and violent crime near places of worship in Recife, Brazil. Do moral communities have a spatial dimension?
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    **Funding:** Edivaldo Constantino das Neves Júnior received support from the Government of Canada, through the program Emerging Leaders in the Americas. **Conflicts of Interest:** The authors declare that they have no conflict of interest.

authors: 
- name: Edivaldo Constantino das Neves Júnior 
  address: University of Sao Paulo - Brazil
  email: edivaldoconstantino@gmail.com

- name: Breno Caldas de Araujo
  address: Federal University of Pernambuco - Brazil
  email: brenocaldasdearaujo@gmail.com

- name: Tatiane Almeida de Menezes
  address: Federal University of Pernambuco - Brazil
  email: tatianedemenezes@gmail.com

- name: Antonio Paez
  address: McMaster University - Canada
  email: paezha@mcmaster.ca

keywords:
- Moral communities
- Crime
- Places of worship
- Point pattern analysis
- Intensity

abstract: |
  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries. With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities. Moral communities, a concept closely related to informal social control, are of interest in countries and regions where formal controls are weak and ineffective. The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship. Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns. Analysis is conducted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets). The results suggest that Catholic places of worship do not project moral communities geographically more than, say, ice cream shops. The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher at short distances than for any of our other referential events; however, at distances between 300 and 500 m the intensity is lower. Finally, the intensity of crime with respect to Spiritist churches, albeit lower at short distances, tends to be significantly higher at distances between 300 and 500 m.

bibliography: bibliography.bib
output: rticles::springer_article
---